---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Package mcst: Overview

<!-- badges: start -->
<!-- badges: end -->

**mcst** is an R package that delivers end-to-end workflows for portfolio risk and return simulations, risk analysis, and stress testing. Key features include:
- Modeling of endogenous and exogenous shocks
- Multiple return distributions (Normal, t, Skewed Generalized T)
- Option for Copula-based dependence structures (Gaussian and t-copula)
- Choice of forward-looking variance estimation for covariance matrix (Sample, EWMA, GARCH)

## Installation

You can install the development version of mcst from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("calebong/mcst")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r load data, include=FALSE}
library(tidyverse)
library(ggplot2)
load( "/Users/Caleb/Documents/Working Directory/mcst/data/returns_data.RData")
load( "/Users/Caleb/Documents/Working Directory/mcst/data/exo_data.RData")

returns_data=returns_data %>% select(week, NVDA, JPM, XOM)
```

## Quick Start:

A sample portfolio of 3 equal-weighted stocks as stocks (endogenous factors) and 2 market-wide (exogenous) factors - SPY and Crude Oil.

```{r pressure, echo = FALSE}
head(returns_data)
head(exo_data)
```

Basic simulation with equal portfolio weights

```{r s1, echo = FALSE, results = 'hide'}
library(mcst)

# Basic simulation with equal weights
result_1 <- portfolio_risk_simulation(
  historical_returns = returns_data
)
```

```{r s1 portfolio distribution, echo = FALSE}
print(result_1$plots$portfolio_diagnostics$portfolio_distribution)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure plot, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
