---
title: "Package mcst: Sample Scenario Illustrations"
subtitle: "Practical Examples of Simulation, Stress Testing, and Results Extraction"
author: "Caleb Ong"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
# In portfolio_risk_simulation_scenarios.Rmd:
vignette: >
  %\VignetteIndexEntry{Portfolio Risk Simulation: Scenario Illustrations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    highlight: tango
    latex_engine: xelatex
    keep_tex: false
geometry: "top=1in, bottom=1in, left=1in, right=1in"
fontsize: 11pt
# mainfont: "Helvetica"
# monofont: "Courier New"
header-includes:
  # - \usepackage{lmodern}
  # - \usepackage[T1]{fontenc}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{xcolor}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{fontspec}
  - \usepackage{pdflscape}
  - \usepackage{float}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  eval      = TRUE,
  fig.align = "center",
  fig.width  = 10,
  fig.height = 6,
  out.width  = "100%"
)
```

```{r libraries, include=FALSE}
library(knitr)
library(kableExtra)
```

\newpage

# Installation of package

The latest development version of the package can be installed using `remotes::install_github("calebong/mcst", dependencies = TRUE)`

```{r install package from githhub,, cache=TRUE, results='hide', message=FALSE, warning=FALSE}
require(remotes)
remotes::install_github("calebong/mcst", dependencies = TRUE)
```

# Data Preparation

## Portfolio Assets

For illustration we use the following packages to download 10 stocks from Yahoo finance at a daily frequency. The returns are then aggregated to weekly returns. These form the **endogenous** assets - the holdings to be simulated.

```{r xts-helper, cache=TRUE}
# simple function to convert class xts to dataframe
xts_to_dataframe <- function(xts_obj, column_names = NULL) {
  dates <- zoo::index(xts_obj)
  df    <- as.data.frame(xts_obj)
  df    <- cbind(date = as.Date(dates), df)
  if (!is.null(column_names)) names(df)[-1] <- column_names
  rownames(df) <- NULL
  df
}
```

```{r endo-data, cache=TRUE}
# load in the libraries to get sample stocks data
library(tidyverse)
library(quantmod)
library(lubridate)

tickers <- c("NVDA","TSM","JPM","AXP","LMT","BA","UAL","XOM","NGG","PG")

out <- map(tickers, function(.x) {
  getSymbols(.x, from = "2010-01-01", to = Sys.Date(),
             src = "yahoo", auto.assign = FALSE) %>%
    xts_to_dataframe() %>%
    dplyr::select(1, 7) %>% #quickly select date and adjusted prices
    rename(Price = 2)
})
names(out) <- tickers

out_wide <- out %>%
  enframe() %>%
  unnest(cols = c(value)) %>%
  pivot_wider(names_from = "name", values_from = "Price")

weekly_returns <- out_wide %>%
  pivot_longer(cols = -date, names_to = "sector", values_to = "price") %>%
  mutate(week = floor_date(date, unit = "week")) %>%
  group_by(sector, week) %>%
  summarise(price_week_end = last(price), .groups = "drop") %>%
  group_by(sector) %>%
  arrange(week) %>%
  mutate(weekly_return = price_week_end / lag(price_week_end) - 1) %>%
  ungroup() %>%
  filter(!is.na(weekly_return)) %>%
  dplyr::select(week, sector, weekly_return) %>%
  pivot_wider(id_cols = week, names_from = sector, values_from = weekly_return)
```

## Exogenous Instruments

SPY (S&P 500 ETF - quick proxy for S&P 500 Index) and Crude Oil futures (`CL=F`) are downloaded separately as
**exogenous** instruments. These carry zero portfolio weight and are used only
as shock transmission channels in the stress-testing step.

```{r exo-data, cache=TRUE}
exo    <- c("SPY", "CL=F")
naming <- c("SPY", "Crude")

out_exo <- map(exo, function(.x) {
  getSymbols(.x, from = "2010-01-01", to = Sys.Date(),
             src = "yahoo", auto.assign = FALSE) %>%
    xts_to_dataframe() %>%
    dplyr::select(1, 7) %>% # quickly select date and adjusted prices
    rename(Price = 2)
})
names(out_exo) <- naming

out_wide_exo <- out_exo %>%
  enframe() %>%
  unnest(cols = c(value)) %>%
  pivot_wider(names_from = "name", values_from = "Price")

weekly_returns_exo <- out_wide_exo %>%
  pivot_longer(cols = -date, names_to = "sector", values_to = "price") %>%
  mutate(week = floor_date(date, unit = "week")) %>%
  group_by(sector, week) %>%
  summarise(price_week_end = last(price), .groups = "drop") %>%
  group_by(sector) %>%
  arrange(week) %>%
  mutate(weekly_return = price_week_end / lag(price_week_end) - 1) %>%
  ungroup() %>%
  filter(!is.na(weekly_return)) %>%
  dplyr::select(week, sector, weekly_return) %>%
  pivot_wider(id_cols = week, names_from = sector, values_from = weekly_return) %>%
  mutate(Crude = ifelse(Crude <= -1, -0.5, Crude))
```

## Combine and Filter

To produce a data frame quickly for illustration, portfolio and exogenous returns are joined on the week key and filtered to start from 2016, giving a cleaner estimation sample.

```{r combine}
comb <- weekly_returns %>%
  full_join(weekly_returns_exo, by = "week") %>%
  mutate(across(where(is.numeric), ~replace_na(., 0))) %>%
  filter(week >= "2016-01-01")

returns_data <- comb %>% dplyr::select(-Crude, -SPY)
exo_data     <- comb %>% dplyr::select(week, Crude, SPY)

cat(sprintf("Portfolio : %d assets, %d weekly obs\n",
            ncol(returns_data) - 1, nrow(returns_data)))
cat(sprintf("Exogenous : %d instruments\n", ncol(exo_data) - 1))

# data structure of input data frames
head(returns_data)
head(exo_data)
```

\newpage

# Scenario 1: Base Case (Gaussian, Equal Weights)

The simplest possible call. No arguments are overridden: the function uses a
multivariate Normal distribution, historical sample covariance, and equal weights
across all ten assets. No stress is applied.


```{r s1-run, cache=TRUE, results="hide"}
library(mcst)

result_1 <- portfolio_risk_simulation(
  historical_returns = returns_data
)
```

## Fitted Marginal Distributions

Each panel shows the empirical return histogram with the fitted Normal density overlaid.

```{r s1-asset-dist, fig.height=9,fig.pos='H',fig.cap="Scenario 1: Fitted Normal marginals vs. empirical returns per asset."}
result_1$plots$portfolio_diagnostics$asset_distributions
```

## Simulated Portfolio Return Distribution

```{r s1-port-dist, fig.height=5,fig.pos='H',fig.cap="Scenario 1: Simulated portfolio return distribution (Gaussian, no stress)."}
result_1$plots$portfolio_diagnostics$portfolio_distribution
```

## Risk Metrics

```{r s1-stats, echo=FALSE}
result_1$portfolio_stats %>%
  kable(booktabs = TRUE, digits = 4,
        col.names = c("Metric","Pre-Stress","Post-Stress"),
        caption = "Scenario 1: Portfolio risk statistics (Gaussian base case).") %>%
  kable_styling(latex_options = c("hold_position","scale_down"), font_size = 9)
```

## Per-Asset Empirical Summary

```{r s1-emp, echo=FALSE}
result_1$asset_empirical_summary %>%
  kable(booktabs = TRUE, digits = 4,
        caption = "Scenario 1: Per-asset empirical return statistics.") %>%
  kable_styling(latex_options = c("hold_position","scale_down"), font_size = 9)
```

\newpage

# Scenario 2: Fat Tails with t-Copula and EWMA Covariance

**Changes from base case:** Multivariate t returns (`rmvt`), t-copula to capture
tail co-dependence, EWMA covariance ($\lambda = 0.97$) so recent observations receive
higher weight, and per-asset degrees-of-freedom calibrated from the data
(`calibration_method = "full"`).


```{r s2-run, cache=TRUE, results="hide"}
result_2 <- portfolio_risk_simulation(
  historical_returns  = returns_data,
  sim_returns_dist    = "rmvt",
  cov_estimation      = "ewma",
  cov_lambda          = 0.97,
  calibrate_params    = TRUE,
  calibration_method  = "full",
  copula              = TRUE,
  copula_type         = "t",
  copula_df           = 4,
  n_sim_returns       = NULL,
  propagation         = TRUE,
  exogenous_returns   = NULL,
  exogenous_shock     = NULL,
  exogenous_volatility_shock = NULL,
  df                  = 4,
  seed                = 123
)
```

## Tail Dependence and Copula Diagnostics

The left panel shows the pairwise lower tail dependence coefficient $\lambda$ - the
probability that two assets move to extreme values simultaneously. The right panel
compares fitted copula correlations to the Kendall $\tau$-implied starting values:
points far from the diagonal indicate that MLE optimisation shifted that pair's
correlation materially from the method-of-moments estimate.

\blandscape

```{r s2-tail, fig.w,fig.height=6,fig.pos='H',fig.cap="Scenario 2: Pairwise tail dependence and copula vs. Kendall baseline."}
result_2$plots$portfolio_diagnostics$tail_dependency
```

\elandscape

## KS Uniformity Test

The KS test checks whether each asset's PIT residuals are Uniform[0,1] after the
t-marginal is fitted. A rejection ($p < 0.05$) flags a misspecified marginal for that
asset.

```{r s2-ks, echo=FALSE}
result_2$ks_test %>%
  kable(booktabs = TRUE, digits = 4,
        caption = "Scenario 2: KS uniformity test on PIT residuals.") %>%
  kable_styling(latex_options = "hold_position", font_size = 9) %>%
  column_spec(4, bold = TRUE)
```


\newpage

# Scenario 3: Exogenous Volatility Shock - SPY Vol x2 (GJR-GARCH)

**Shock:** SPY volatility doubles ($k_{SPY} = 2$); Crude is unchanged ($k_{Crude} = 1$).
No return shock is applied, so **portfolio means are unaffected** - only the
covariance widens via the Law of Total Variance addon. GJR-GARCH estimates per-asset
forward volatility, capturing volatility clustering and the leverage effect.

```{r s3-shocks}
exogenous_volatility_shock <- list(SPY = 2, Crude = 1)
```

```{r s3-run, cache=TRUE, results="hide"}
result_3 <- portfolio_risk_simulation(
  historical_returns          = returns_data,
  sim_returns_dist            = "rmvt",
  cov_estimation              = "garch",
  calibrate_params            = TRUE,
  calibration_method          = "full",
  copula                      = TRUE,
  copula_type                 = "t",
  copula_df                   = 4,
  n_sim_returns               = 10000,
  propagation                 = TRUE,
  exogenous_returns           = exo_data,
  exogenous_shock             = NULL,
  exogenous_volatility_shock  = exogenous_volatility_shock,
  df                          = 4,
  seed                        = 123
)
```

## GJR-GARCH Conditional Volatility Paths

Each panel shows: salmon bars = raw returns; blue line = in-sample conditional
$\hat{\sigma}_t$; red dashed = long-run unconditional $\bar{\sigma}$; green diamond
= 1-step-ahead forecast. Assets where the forecast sits well above $\bar{\sigma}$
are in an elevated-volatility regime.

\blandscape

```{r s3-garch, fig.height=10,fig.pos='H',fig.cap="Scenario 3: GJR-GARCH conditional volatility paths per asset."}
result_3$plots$portfolio_diagnostics$time_series
```

\elandscape

## Exogenous Propagation Channel

The beta $\beta_{pe} = \text{Cov}(X_p, X_e) / \text{Var}(X_e)$ quantifies the
transmission of SPY volatility to each portfolio asset. With a pure vol shock (no
return shock), these betas determine the size of the LTV covariance addon - not
any mean shift.

\blandscape


```{r s3-prop, fig.height=5,fig.pos='H',fig.cap="Scenario 3: SPY propagation betas per portfolio asset."}
result_3$stressed_plots$exogenous_propagation
```


\elandscape

\blandscape
## Portfolio Distribution: Pre vs. Post Stress


```{r s3-dist, fig.height=5,fig.pos='H',fig.cap="Scenario 3: Portfolio return distribution, base vs. SPY vol x2."}
result_3$stressed_plots$portfolio_distribution_comparison
```

\elandscape


## Stress Impact Summary

```{r s3-impact, echo=FALSE}
result_3$stress_impact_summary %>%
  kable(booktabs = TRUE, digits = 4,
        caption = "Scenario 3: Risk metric changes from SPY volatility doubling.") %>%
  kable_styling(latex_options = "hold_position", font_size = 9)
```

\newpage

# Scenario 4: Oil Spike and Market Selloff (Exogenous Return and Vol Shocks)

**Shock:** SPY falls 5% ($\delta_{SPY} = -0.05$) with vol doubled; Crude surges 20%
($\delta_{Crude} = +0.20$) with vol doubled. This may simulate a stagflationary stress scenario: energy
costs surge while equity markets sell off simultaneously.

Both return shocks enter the Schur complement together. The shocked block is
$s = \{SPY, Crude\}$ and all portfolio asset means are updated by
$\mu_{p|s} = \mu_p + \Sigma_{pe}\Sigma_{ee}^{-1}\delta_e$. Energy-sensitive
holdings (XOM) received a net positive effect, whereby a positive effect from a high beta to Crude Oil is partially offset by a negative effect from SPY declines. Broad-market holdings (JPM, AXP, NVDA) are dominated by the SPY decline.

```{r s4-shocks}
exogenous_shock            <- list(SPY = -0.05, Crude = 0.20)
exogenous_volatility_shock <- list(SPY = 2,     Crude = 2)
```

```{r s4-run, cache=TRUE, results="hide"}
result_4 <- portfolio_risk_simulation(
  historical_returns          = returns_data,
  sim_returns_dist            = "rmvt",
  cov_estimation              = "garch",
  calibrate_params            = TRUE,
  calibration_method          = "full",
  copula                      = TRUE,
  copula_type                 = "t",
  copula_df                   = 4,
  n_sim_returns               = NULL,
  propagation                 = TRUE,
  exogenous_returns           = exo_data,
  exogenous_shock             = exogenous_shock,
  exogenous_volatility_shock  = exogenous_volatility_shock,
  df                          = 4,
  seed                        = 123
)
```

## Propagation of the Compound Exogenous Shock

\blandscape

```{r s4-prop, fig.height=5,fig.pos='H',fig.cap="Scenario 4: Conditional mean shifts from SPY -5pct and Crude +20pct."}
result_4$stressed_plots$exogenous_propagation
```

\elandscape

\blandscape

## Portfolio Distribution Comparison

```{r s4-dist, fig.height=5,fig.pos='H', fig.cap="Scenario 4: Portfolio distribution under oil spike and market selloff."}
result_4$stressed_plots$portfolio_distribution_comparison
```
\elandscape

## Stress Impact Summary

```{r s4-impact, echo=FALSE}
result_4$stress_impact_summary %>%
  kable(booktabs = TRUE, digits = 4,
        caption = "Scenario 4: Risk metrics under stagflationary shock.") %>%
  kable_styling(latex_options = "hold_position", font_size = 9)
```

## Exogenous Historical Correlations

This table explains the propagation magnitudes above. Assets with high SPY correlation
(JPM, AXP, NVDA) absorb the largest mean shift from the SPY leg; XOM's positive Crude
correlation provides a partial offset.

```{r s4-corr, echo=FALSE}
result_4$exogenous_historical_correlation %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Scenario 4: Historical correlations, portfolio vs. SPY and Crude.") %>%
  kable_styling(latex_options = "hold_position", font_size = 9)
```

\newpage

# Scenario 5: Semiconductor Selloff with SPY Offset (Mixed Endo and Exog)

**Shock:** NVDA and TSM each fall 5% with vol multiplied 1.5x (endogenous return
and vol shocks). Simultaneously, SPY rises 2.5% with vol 1.25x (exogenous return
and vol shock). This captures a sector-specific stress ie regulatory action or
export controls on semiconductors that occurs against a mild broader market rally.

The Schur complement operates on a combined shocked block $s = \{NVDA, TSM, SPY\}$.
Assets correlated with the semiconductor pair receive a negative propagated mean
shift; the SPY tailwind partially offsets this for holdings with high market beta.

```{r s5-shocks}
asset_shock                <- list(NVDA = -0.05, TSM = -0.05)
volatility_shock           <- list(NVDA = 1.5,   TSM = 1.5)
exogenous_shock            <- list(SPY = 0.025)
exogenous_volatility_shock <- list(SPY = 1.25)
```

```{r s5-run, cache=TRUE, results="hide"}
result_5 <- portfolio_risk_simulation(
  historical_returns          = returns_data,
  sim_returns_dist            = "rmvt",
  cov_estimation              = "garch",
  calibrate_params            = TRUE,
  calibration_method          = "full",
  copula                      = TRUE,
  copula_type                 = "t",
  copula_df                   = 4,
  n_sim_returns               = NULL,
  propagation                 = TRUE,
  asset_shock                 = asset_shock,
  volatility_shock            = volatility_shock,
  exogenous_returns           = exo_data,
  exogenous_shock             = exogenous_shock,
  exogenous_volatility_shock  = exogenous_volatility_shock,
  df                          = 4,
  seed                        = 123
)
```

## Mean Shift Decomposition (Waterfall)

The stacked bar chart decomposes each asset's total conditional mean shift into three
components: (1) **direct endogenous shock** (non-zero only for NVDA and TSM), (2)
**propagated from endo** (Schur from the NVDA/TSM block), and (3) **propagated from
exog** (Schur from the SPY return shock). The diamond marker shows the net total.

\blandscape

```{r s5-mixed,fig.pos='H', fig.height=6,fig.cap="Scenario 5: Mean shift decomposition - direct endo, propagated endo, propagated exog."}
result_5$stressed_plots$mixed_shock_effects
```

\elandscape

## Portfolio Distribution Comparison

Despite the SPY tailwind, the portfolio distribution shifts left because NVDA and TSM
are large equal-weight components with high intra-sector correlations that propagate
their decline across correlated holdings.

\blandscape

```{r s5-dist, fig.height=5,fig.cap="Scenario 5: Portfolio distribution - semiconductor selloff with SPY rally."}
result_5$stressed_plots$portfolio_distribution_comparison
```
\elandscape

## Stress Impact Summary

```{r s5-impact, echo=FALSE}
result_5$stress_impact_summary %>%
  kable(booktabs = TRUE, digits = 4,
        caption = "Scenario 5: Risk metrics under semiconductor-specific stress.") %>%
  kable_styling(latex_options = "hold_position", font_size = 9)
```

\newpage

# Accessing Simulation Draws Directly

The raw per-asset simulated draws are stored as $n_{\text{sim}} \times n_{\text{assets}}$
data frames - accessible without any aggregation to portfolio level. This enables
custom downstream analysis such as per-asset VaR.

```{r draws-access}
draws_pre  <- result_5$asset_draws
draws_post <- result_5$asset_draws_stress

cat("Pre-stress  draws:", nrow(draws_pre),  "rows x", ncol(draws_pre),  "cols\n")
cat("Post-stress draws:", nrow(draws_post), "rows x", ncol(draws_post), "cols\n")
cat("Columns:", paste(names(draws_pre), collapse = ", "), "\n")
```

## NVDA: Pre vs. Post Stress Simulated Distribution

```{r draws-plot, fig.height=4, fig.cap="NVDA simulated return distribution, pre and post Scenario 5 stress."}
nvda_comp <- data.frame(
  Return   = c(draws_pre$NVDA, draws_post$NVDA),
  Scenario = rep(c("Pre-Stress","Post-Stress"), each = nrow(draws_pre))
)

means_df <- nvda_comp %>%
  group_by(Scenario) %>%
  summarise(m = mean(Return * 100), .groups = "drop")

ggplot(nvda_comp, aes(x = Return * 100, fill = Scenario)) +
  geom_density(alpha = 0.45, colour = NA) +
  geom_vline(data = means_df, aes(xintercept = m, colour = Scenario),
             linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(values   = c("Pre-Stress" = "steelblue","Post-Stress" = "coral")) +
  scale_colour_manual(values = c("Pre-Stress" = "steelblue","Post-Stress" = "coral")) +
  labs(title    = "NVDA: Simulated Return Distribution, Pre vs. Post Stress",
       subtitle = "Scenario 5: NVDA direct shock -5%, vol x1.5 | SPY +2.5%, vol x1.25",
       x = "Weekly Return (%)", y = "Density", fill = NULL, colour = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

\newpage

# Cross-Scenario Comparison

```{r compare, echo=FALSE}
get_stat <- function(res, metric) {
  res$portfolio_stats$Pre_Stress[res$portfolio_stats$Metric == metric]
}

comparison <- data.frame(
  Scenario = c(
    "S1: Gaussian, sample cov (no stress)",
    "S2: t + EWMA + t-copula (no stress)",
    "S3: t + GARCH, SPY vol x2",
    "S4: t + GARCH, SPY -5pct / Crude +20pct",
    "S5: t + GARCH, NVDA/TSM -5pct + SPY +2.5pct"
  ),
  Mean   = round(c(get_stat(result_1,"Mean (%)"),     get_stat(result_2,"Mean (%)"),
                   get_stat(result_3,"Mean (%)"),     get_stat(result_4,"Mean (%)"),
                   get_stat(result_5,"Mean (%)")), 4),
  Vol    = round(c(get_stat(result_1,"Volatility (%)"),get_stat(result_2,"Volatility (%)"),
                   get_stat(result_3,"Volatility (%)"),get_stat(result_4,"Volatility (%)"),
                   get_stat(result_5,"Volatility (%)")), 4),
  VaR95  = round(c(get_stat(result_1,"VaR 95% (%)"),  get_stat(result_2,"VaR 95% (%)"),
                   get_stat(result_3,"VaR 95% (%)"),  get_stat(result_4,"VaR 95% (%)"),
                   get_stat(result_5,"VaR 95% (%)")), 4),
  CVaR95 = round(c(get_stat(result_1,"CVaR 95% (%)"), get_stat(result_2,"CVaR 95% (%)"),
                   get_stat(result_3,"CVaR 95% (%)"), get_stat(result_4,"CVaR 95% (%)"),
                   get_stat(result_5,"CVaR 95% (%)")), 4),
  stringsAsFactors = FALSE
)

kable(comparison, booktabs = TRUE,
      col.names = c("Scenario","Mean (%)","Vol (%)","VaR 95% (%)","CVaR 95% (%)"),
      caption = "Pre-stress portfolio risk statistics across all five scenarios.") %>%
  kable_styling(latex_options = c("hold_position","scale_down"), font_size = 9)
```

*Scenarios 3--5 show the pre-stress baseline. The incremental stress impact is
captured in each scenario's stress impact summary table above.*

\newpage

# Quick Reference: Result Object Fields

```{r ref-table, echo=FALSE}
ref <- data.frame(
  Field = c(
    "result$portfolio_stats",
    "result$asset_draws",
    "result$asset_draws_stress",
    "result$asset_empirical_summary",
    "result$exogenous_historical_correlation",
    "result$exogenous_asset_empirical_summary",
    "result$ks_test",
    "result$stress_impact_summary",
    "result$stressed_covariance",
    "result$stressed_means",
    "result$final_covariance",
    "result$copula_correlation",
    "result$tail_dependence",
    "result$calibrated_params",
    "result$plots$...$asset_distributions",
    "result$plots$...$portfolio_distribution",
    "result$plots$...$tail_dependency",
    "result$plots$...$time_series",
    "result$stressed_plots$portfolio_distribution_comparison",
    "result$stressed_plots$exogenous_propagation",
    "result$stressed_plots$mixed_shock_effects"
  ),
  Description = c(
    "14-row data frame: risk metrics pre and post stress",
    "n_sim x n_assets data frame of pre-stress per-asset draws",
    "n_sim x n_assets data frame of post-stress per-asset draws",
    "Per-asset empirical mean, SD, skewness, kurtosis, min, max",
    "Portfolio x exogenous Pearson correlations (Asset + one col per instrument)",
    "Exogenous instrument empirical summary: mean, SD, skewness, kurtosis",
    "KS test results: Asset, KS_stat, p_value, Flag",
    "7-row data frame: Metric, Pre_Stress, Post_Stress, Change",
    "n x n stressed covariance matrix after shock application",
    "Named vector of stressed conditional means per asset",
    "n x n final forward-looking covariance matrix",
    "n x n fitted copula correlation matrix",
    "n x n tail dependence matrix (t-copula only)",
    "Calibrated shape parameters per asset",
    "Histogram + fitted density overlay, one panel per asset",
    "Simulated portfolio return distribution (histogram + KDE)",
    "Tail dependence bar chart + copula vs. Kendall tau scatter",
    "GJR-GARCH conditional vol time series (GARCH cov_estimation only)",
    "Overlaid pre vs. post stress portfolio return distributions",
    "Exogenous beta and propagation channel plots (Scenario B)",
    "Waterfall decomposition: direct + propagated endo + exog (Scenario C)"
  ),
  stringsAsFactors = FALSE
)

kable(ref, booktabs = TRUE,
      col.names = c("Field","Description"),
      caption = "Quick reference: result object fields and plot accessors.") %>%
  kable_styling(latex_options = c("hold_position","scale_down","repeat_header"),
                font_size = 8) %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "6.5cm")
```

\newpage

# Sample Function Output: Console Printout

The `portfolio_risk_simulation` function produces a detailed diagnostic
console printout during execution. This section captures a representative
example from Scenario 4 (oil spike and market selloff) to illustrate what
the output looks like, covering data preprocessing, copula fitting,
stress-testing steps, and the final stress impact summary.

```{r printout-capture, results="asis", echo=FALSE}
# Capture the full console printout by re-running with capture.output().
# cache=FALSE is intentional here: we always want a fresh capture.
printout <- capture.output({
  invisible(portfolio_risk_simulation(
    historical_returns          = returns_data,
    sim_returns_dist            = "rmvt",
    cov_estimation              = "garch",
    calibrate_params            = TRUE,
    calibration_method          = "full",
    copula                      = TRUE,
    copula_type                 = "t",
    copula_df                   = 4,
    n_sim_returns               = NULL,
    propagation                 = TRUE,
    exogenous_returns           = exo_data,
    exogenous_shock             = list(SPY = -0.05, Crude = 0.20),
    exogenous_volatility_shock  = list(SPY = 2, Crude = 2),
    df                          = 4,
    seed                        = 123
  ))
})

cat(paste(printout, collapse = "\n"))

# # Emit landscape environment around a verbatim block
# cat("\
# 
# ")
# cat("\\begin{verbatim}
# ")
# cat(paste(printout, collapse = "\n"))
# cat("
# \\end{verbatim}
# 
# ")
# cat("\
# ")
```

